---
title: "HW6 "
author: "Jake Grossman"
editor: visual
format:
  html:
    embed-resources: true
execute:
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(tidytext)
library(schrute)
```

# The Office

Let's continue to work with *The Office* text from the [schrute](https://cran.r-project.org/web/packages/schrute/index.html) package.

```{r}
library(schrute)
library(tidyverse)
library(tidytext)
library(lexicon)
library(kableExtra)

```

# Sentiment Analysis

```{r}
office <- schrute::theoffice

## Make sure air_date is a Date
office <- office %>%
  mutate(air_date = as.Date(air_date))

bing <- get_sentiments("bing")
bing

# Episode-level metadata (one row per episode)
episode_lite <- office %>%
  group_by(season, episode) %>%
  summarise(
    episode_name = first(episode_name),
    imdb_rating  = first(imdb_rating),
    air_date     = first(air_date),
    .groups = "drop"
  ) %>%
  arrange(air_date, season, episode)
episode_lite
```

### Over Time

Calculate the sentiment score (positive - negative) for each episode, and plot this variable over time.

```{r}
episode_sent <- office %>%
  mutate(text = str_to_lower(text)) %>%
  unnest_tokens(word, text) %>%
  inner_join(bing, by = "word") %>%
  count(season, episode, sentiment, name = "n") %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sent_score = positive - negative) %>%
  right_join(episode_meta, by = c("season","episode")) %>%
  replace_na(list(sent_score = 0, positive = 0, negative = 0))

episode_sent
# Preportion for episodes (and testing out kableExtra)
episode_sent <- episode_sent %>% mutate(pos_prop = positive/(negative+positive))


episode_kable <- episode_sent %>%
  select(season, episode_name, imdb_rating, pos_prop) %>%
  arrange(desc(pos_prop)) %>% slice_head(n=10)

episode_kable %>% kable() %>% kable_minimal() %>%
  collapse_rows(columns = 1, valign = "top") %>% 
    kable_styling(
    bootstrap_options = c("striped", "hover"),
    position = "left",
    font_size = 14,
    fixed_thead = T)

ggplot(episode_sent_naive, aes(x = air_date, y = sent_score)) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.7) +
  labs(
    title = "Episode sentiment over time",
    x = "Air date",# need Air date due to episodes having the same number per-season
    y = "Sentiment score (positive - negative)"
  ) +
  theme_minimal()

```

### By Rating

Create a scatterplot of the `imdb_rating` versus the sentiment score (positive - negative), where the case is again an episode. Decide whether a linear model is a good fit to this relationship (optional, if you have had modeling classes, what model looks best?). Comment on what you find.

```{r}
df_rating_naive <- episode_sent %>% filter(!is.na(imdb_rating))

ggplot(df_rating_naive, aes(x = imdb_rating, y = sent_score)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(method = "loess", se = FALSE, linetype = 2) +
  labs(
    title = "IMDb rating vs sentiment (naive)",
    x = "IMDb rating",
    y = "Sentiment score (positive - negative)"
  ) +
  theme_minimal()

m_lm <- lm(sent_score ~ imdb_rating, data = df_rating_naive)
summary(m_lm)

m_poly2 <- lm(sent_score ~ poly(imdb_rating, 2), data = df_rating_naive)
AIC(m_lm, m_poly2)


```

The Scatter plot shows a slight positive relation between the IMDb rating and the sentiment score. it shows that fro every \~13 more positive words the IMDb score goes up a point, also the R\^2 value is \~ 0.044 so the rating only explains about 4% of the variability. And the non linear model is slightly better due to a low AIC however I don't think that the poly2 is the best option for this case.

### Sentiment Negation

Identify at least a few negations that change the sentiment values based on the text. Use this to recreate the graphs above (or do this first, then create the graphs).

```{r}
# Define negation words
negation_words <- c("not", "no", "never", "without", "don't", "doesn't", 
                    "didn't", "won't", "can't", "cannot", "hardly", "barely")

# Tokenize into bigrams to capture negation context
office_bigrams <- theoffice %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ")

# Identify bigrams where word1 is a negation
negated_words <- office_bigrams %>%
  filter(word1 %in% negation_words) %>%
  inner_join(get_sentiments("afinn"), by = c("word2" = "word")) %>%
  mutate(value = -value)  # flip the sentiment score

# Non-negated sentiment (standard)
normal_sentiment <- office_bigrams %>%
  filter(!word1 %in% negation_words) %>%
  inner_join(get_sentiments("afinn"), by = c("word2" = "word"))

# Combine both
adjusted_sentiment <- bind_rows(normal_sentiment, negated_words)

# Summarise by episode
episode_sent_negation <- adjusted_sentiment %>%
  group_by(season, episode, air_date) %>%
  summarise(sent_score = sum(value), .groups = "drop")

# Plot
ggplot(episode_sent_negation, aes(x = air_date, y = sent_score)) +
  geom_col(width = 0.4) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  labs(
    title = "Episode sentiment over time (negation-adjusted)",
    x = "Episode",
    y = "Sentiment score (positive - negative)"
  )

episode_kable_neg <- episode_sent_negation %>%
  select(season, episode, sent_score) %>%
  arrange(desc(sent_score)) %>% slice_head(n=10)

episode_kable_neg %>% kable() %>% kable_minimal() %>%
  collapse_rows(columns = 1, valign = "top") %>% 
    kable_styling(
    bootstrap_options = c("striped", "hover"),
    position = "left",
    font_size = 14,
    fixed_thead = T)
```
