---
title: "HW6 "
author: "Jake Grossman"
editor: visual
format:
  html:
    embed-resources: true
execute:
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(tidytext)
library(schrute)
```

# The Office

Let's continue to work with *The Office* text from the [schrute](https://cran.r-project.org/web/packages/schrute/index.html) package.

```{r}
library(schrute)
library(tidyverse)
library(tidytext)
library(lexicon)
library(kableExtra)

```

# Sentiment Analysis

```{r}
office <- schrute::theoffice %>%
  mutate(air_date = as.Date(air_date))

bing <- get_sentiments("bing")

# Episode-level metadata (one row per episode)
episode_meta <- office %>%
  group_by(season, episode) %>%
  summarise(
    episode_name = first(episode_name),
    imdb_rating  = first(imdb_rating),
    air_date     = first(air_date),
    .groups = "drop"
  ) %>%
  arrange(air_date, season, episode)

episode_meta
```

### Over Time

Calculate the sentiment score (positive - negative) for each episode, and plot this variable over time.

```{r}
episode_sent <- office %>%
  mutate(text = str_to_lower(text)) %>%
  unnest_tokens(word, text) %>%
  inner_join(bing, by = "word") %>%
  count(season, episode, sentiment, name = "n") %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sent_score = positive - negative) %>%
  right_join(episode_meta, by = c("season", "episode")) %>%
  replace_na(list(sent_score = 0, positive = 0, negative = 0)) %>%
  mutate(pos_prop = positive / (positive + negative))

episode_sent

episode_kable <- episode_sent %>%
  select(season, episode_name, imdb_rating, pos_prop) %>%
  arrange(desc(pos_prop)) %>%
  slice_head(n = 10)

episode_kable %>%
  kable() %>%
  kable_minimal() %>%
  collapse_rows(columns = 1, valign = "top") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    position = "left",
    font_size = 14,
    fixed_thead = TRUE
  )

ggplot(episode_sent, aes(x = air_date, y = sent_score)) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.7) +
  labs(
    title = "Episode sentiment over time",
    x = "Air date",
    y = "Sentiment score (positive - negative)"
  ) +
  theme_minimal()
```

\*Comment: I have also started trying to implement kable and kableExtra in my work, so I really liked your use of kable to present the top episodes. I find using these tables allows results to be displayed in a much more clean, readable, and organized way. The styling choices you make keep things simple, while also making sure to highlight the necessary information.

### By Rating

Create a scatterplot of the `imdb_rating` versus the sentiment score (positive - negative), where the case is again an episode. Decide whether a linear model is a good fit to this relationship (optional, if you have had modeling classes, what model looks best?). Comment on what you find.

```{r}
df_rating_naive <- episode_sent %>% filter(!is.na(imdb_rating))

ggplot(df_rating_naive, aes(x = imdb_rating, y = sent_score)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(method = "loess", se = FALSE, linetype = 2) +
  labs(
    title = "IMDb rating vs sentiment (naive)",
    x = "IMDb rating",
    y = "Sentiment score (positive - negative)"
  ) +
  theme_minimal()

m_lm <- lm(sent_score ~ imdb_rating, data = df_rating_naive)
summary(m_lm)

m_poly2 <- lm(sent_score ~ poly(imdb_rating, 2), data = df_rating_naive)
AIC(m_lm, m_poly2)


```

The Scatter plot shows a slight positive relation between the IMDb rating and the sentiment score. it shows that fro every \~13 more positive words the IMDb score goes up a point, also the R\^2 value is \~ 0.044 so the rating only explains about 4% of the variability. And the non linear model is slightly better due to a low AIC however I don't think that the poly2 is the best option for this case.

\*Comment: As a non-statistics major, I really appreciate and enjoy your explanation here of the relationship between IMDBb rating and sentiment. I find the explanation very clear and informative, especially how you reference both the significance and meaning of the slope and R\^2 values. I like how you tied the statistical results back to the overall takeaway of the graph, making it easy to interpret the findings.

### Sentiment Negation

Identify at least a few negations that change the sentiment values based on the text. Use this to recreate the graphs above (or do this first, then create the graphs).

```{r}
# Define negation words
negation_words <- c("not", "no", "never", "without", "don't", "doesn't", 
                    "didn't", "won't", "can't", "cannot", "hardly", "barely")

# Tokenize into bigrams
office_bigrams <- office %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ")

# Negated sentiment
negated_words <- office_bigrams %>%
  filter(word1 %in% negation_words) %>%
  inner_join(get_sentiments("afinn"), by = c("word2" = "word")) %>%
  mutate(value = -value)

# Non-negated sentiment
normal_sentiment <- office_bigrams %>%
  filter(!word1 %in% negation_words) %>%
  inner_join(get_sentiments("afinn"), by = c("word2" = "word"))

# Combine
adjusted_sentiment <- bind_rows(normal_sentiment, negated_words)

# Summarise by episode, then re-attach episode metadata
episode_sent_negation <- adjusted_sentiment %>%
  group_by(season, episode) %>%
  summarise(sent_score = sum(value), .groups = "drop") %>%
  left_join(episode_meta, by = c("season", "episode"))

ggplot(episode_sent_negation, aes(x = air_date, y = sent_score)) +
  geom_col(width = 0.4) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  labs(
    title = "Episode sentiment over time (negation-adjusted)",
    x = "Air date",
    y = "Sentiment score"
  ) +
  theme_minimal()

episode_kable_neg <- episode_sent_negation %>%
  select(season, episode, sent_score) %>%
  arrange(desc(sent_score)) %>%
  slice_head(n = 10)

episode_kable_neg %>%
  kable() %>%
  kable_minimal() %>%
  collapse_rows(columns = 1, valign = "top") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    position = "left",
    font_size = 14,
    fixed_thead = TRUE
  )
```

\*Comment: I found your use of the afinn sentiment lexicon in your sentiment negation really interesting. Before looking at your work, I was pretty unfamiliar with it, but after figuring out what it is and how you utilized it to adjust sentiment values, I really was able to see and understand its use case.

\*Suggestion: One suggestion I might have, which is something I'm looking to work on myself, is adding a bit more detail to plot labels or captions. I think doing this can help guide the viewer on what they are supposed to take away from each graph/visual, especially if the viewer isn't getting direct access to your code or comments. For example, in your plot of episode sentiment over time, you could briefly clarify what a higher or lower sentiment score represents.
